Are populations like a circuit? Comparing isolation by resistance to a new coalescent-based method.

### 2019-04-18 review-responses.tex

\begin{point}{\revref}
    Do you actually use this tensor product for the movement
    rates anywhere? And seems Eq.\ 7 (which has no tensor product) is all you
    need.
\end{point}

\reply{
    We included this because we think it's a nice compact way to write, code, 
    and think about the equation,
    and the tensor product equation is actually the one used in the R code.
    It is correct that the equations are equivalent.
}

\begin{point}{Line 115:}
    Not all $ZC$ skew-symmetric would translate to valid generators 
    $G + Z$. The diagonal elements of G have another condition (they have to be minus
    the sum of all other jumping rates). Would be interesting to explore this a
    bit more -- overall what the dimensionality this degeneracy is (implicit
    function theorem), and also if there are simple toy examples where they pop
    up.
\end{point}

\reply{
    Thank you for catching this.
    You are correct that in addition to $ZC$ being skew-symmetric,
    the rows of $Z$ must also sum to zero, which we now say. {\revref}
    Further exploration of the constraints would be interesting,
    but we think is beyond the scope of this paper.
}

@@

    This paper implements such a method and conducts comparative analyses on a
    number of simulated data sets. It seems that in general, this method and the
    resistance-based approaches (as exemplified by EEMS) produce somewhat com-
    parable results so long as migration rates are symmetric. The largest differences
    emerge when gene flow is asymmetric, because this fundamentally violates the
    resistance distance model used by EEMS. Here, the authors show convincingly
    that inferring migration rates using the resistance distance approximation can
    be wildly inaccurate compared to the same estimates obtained using coalescent
    methods. To the extent that gene flow is more likely to be asymmetric in many
    natural populations, this seems like an important finding.

    My overall impression of this work is fairly positive. The paper is very clearly
    written, and the idea of replacing IBR with a method that is more faithful to
    the underlying biology is appealing. The methods section (as well as appendices)
    feature a very nice and readable summary of the mathematical innovations of
    this work, as well as some interesting discussion of issues of identifiability and
    power. Similarly, discussion section in very well done and really strengthens
    the manuscript. It includes a thorough survey of the related work, as well as a
    forthright discussion of advantages and disadvantages of the proposed method,
    and it raises a number of interesting questions and directions for future research.
    Overall, the paper was a pleasure to read.
\end{quote}

Thanks very much for the kind words.
One minor comment: resistance distances are not strictly based in a population genetics model,
but we think that in practice it is thought of as (an approximation to) such a model
-- for instance, it is common to see the explanation that it
``integrates over all possible paths that the organisms might take''.
We think it is clear that \citet{mcrae2006isolation} saw resistance distance as an approximation
to the more-accurate coalescence models; we're looking at the effects of this approximation.
We hope this point comes through in the paper.

@@

\begin{quote}
    I find the main deficiency to be the lack of any analysis of real data. All
    experiments are carried out on simulated data only. This is a puzzling omission.
    The authors went to the trouble of creating an R package (which is good), so
    clearly they intend for the method to be used in practice. Personally, I would
    hesitate to publish the results obtained from a method which had not previously
    been verified to produce reasonable answers in real data. This really feels like
    something the authors ought to have included. It is not necessary that the data
    be novel -- reanalyzing (say) the African elephant dataset in the EEMS paper
    would provide for an interesting comparison (even if it cannot be done at an
    equivalent spatial resolution) and would also go long way towards reassuring
    practitioners.
\end{quote}

We've now added analysis of a real dataset, from \citet{geraldes2014landscape},
and are quite pleased with the results!
This has also given us a chance to spend more time talking through how to interpret the results.
(We didn't get to say this in the paper,
but initially were worried by the southward bias in gene flow for the inland poplar species,
since we would expect the opposite if poplars had expanded to the north following the glaciers;
but then we learned that this species likely expanded from Alaksa, rather than from the south.)

@@

\begin{point}{}
    ``Bayesian inference of movement rates'': In their model the authors assume
    that many entries of the migration rate matrix are zero, and place an
    exponential prior on the rest. Is it safe to assume that we will always
    want to constrain rates between discontiguous cells to zero? Also, given
    the valid concerns that they raise elsewhere about identifiability and
    power, I wonder if the authors considered placing a sparse prior on the
    entire G matrix, enabling the model to adapt to the amount of signal while
    shrinking other migration rates to zero. This might have the effect of
    making the resulting non-zero estimates more stable. (I realize that this
    raises potential problems for inference if the underlying graph becomes
    disconnected.)
\end{point}

\reply{
    Whether or not it is reasonable to assume there is no long distance migration
    depends on the population under study. 
    Putting a sparse prior on all of $G$ is a good idea for future improvement.
    % although it would probably be appropriate for the prior probability of a value
    % to be non-zero to be much higher for adjacent demes as compared to non-adjacent ones, 
    % which would also likely prevent the graph from becoming disconnected very often.
    (Since disconnecting the graph would result in some infinite coalescence times, 
    such proposals would be rejected by the MCMC algorithm.)
}

@@

\begin{point}{}
    The discussion of ill-conditioning is welcome. From experience, it often
    feels like coalescent-based methods have little power to accurately
    estimate migration unless the rates are large. It might be interesting to
    shift emphasis from estimation to the slightly easier problem of testing:
    how good are these methods at distinguishing between a given null and an
    alternative hypothesis (say, barrier vs.\ no barrier) for the rate matrix $G$?
    I think the gain of using a coalescent model might be more apparent in this case.
\end{point}

\reply{
    This is a good point; we've added it to the discussion
    (and credited you with it). \revref
}

\begin{point}{}
    This method is computationally limited by equation (7), which has a
    complexity of $O(d^6)$ for the product Markov chain on d demes mentioned a
    few lines below. At least, that is the cost for solving (7) exactly -- but
    given the noisiness of the matrix C, there may be little point in doing
    that anyways. I wonder if the authors have thought at all about obtaining
    an approximate solution to (7), where the error in approximation is chosen
    to in some way to balance out the statistical error inherent in C / D. It
    may also be possible to leverage the fact that the solution G is sparse
    here.  (This is more of a suggestion for future research rather than a
    suggestion for revision.)
\end{point}

\reply{
    Thanks for the suggestions.
    Solving (7) directly for $G$ and $\gamma$ with a noisy $C$ is extremely difficult
    even in cases small enough to just use least squares; 
    it typically resulted in some negative values unless there is very little noise in $C$.
    In order to solve for $C$ from $G$ and $\gamma$, 
    we experimented with the Jacobi method. 
    While it seemed to converge fairly reliably, 
    at the grid sizes where it was computationally advantageous to use it over solving exactly, 
    it would still take a very long time for the MCMC to converge
    due to the large number of parameters.
    % Devoloping a way to reduce the number of parameters in a way similar to EEMS
    % while still allowing asymmetric migration is a good direction for future implementations.
}

@@

\begin{point}{\revref}
     Do the authors have any other guidance on how to discretize the landscape?
\end{point}

\reply{
    This is a large, thorny issue,
    closely related to the question of how to translate discrete-population models
    to continuous geography.
    We've added a bit more \revref,
    as well as some more discussion of this in the section on poplar data.
    \llname{ll:discretization}
}

@@

\begin{point}{}
    Does the issue of having negative entries of G (lines 112-117, Appendix A)
    arise during the course of MCMC sampling, and if so, how is it dealt with?
\end{point}

\reply{
    No. The proposal distribution includes taking the absolute value of parameters
    (a standard method in MCMC to preserve reversibility).
    We've added this note to the MCMC methods section. \revref
}

@@

\reply{
    We think this is correct (and reads reasonably well?) as it stands.
}

\reviewersection{3}

\begin{quote}
    This is an interesting paper on a topic that should be of wide interest to
    readers of MER.  The main results are useful: 
    \begin{enumerate}
        \item      landscape genetics models based on resistance theory can be
            misleading if gene flow is asymmetrical
        \item      coalescent-based models deal better with asymmetry but are
            not a panacea for a variety of reasons.
    \end{enumerate}
    On the other hand, I found myself with quite a few questions after reading
    the manuscript.  Raising new questions is not bad in and of itself, but
    most of these were questions that could/should have been clarified in the
    text.  As a consequence, the manuscript is not as effective or useful as it
    might be.  The problem is not in the writing itself, which generally is
    careful and precise.  Instead, the main problem involves translating a very
    technical exercise into something that is accessible to typical readers of
    the journal.  Some examples are described below.

    Although the theme of the paper is relevant to the journal, I think most
    readers of MER would find the manuscript hard going in many places.  For
    example, the text at 102-109 and 136-146 would benefit from translation
    into insights more meaningful to biologists.  The authors do a good job of
    this in discussing results for asymmetric gene flow at lines 296-299. % \llname{ll:dense3}
\end{quote}

Thanks for the encouragement.
We've tried to improve accessibility in a number of places,
in particular by adding a ``flag'' at the top of the methods {\llname{ll:flagging}},
and improving the two bits of text referred to,
at {\llname{ll:dense1}} and {\llname{ll:dense2}}.

@@

\begin{quote}
    The reader also needs to understand the biological consequences of various
    assumptions.  Examples  where this could be improved include:
\end{quote}

\begin{point}{\revref}
      if the genome is selectively neutral ...
\end{point}

\reply{
    Reworded.
}

@@

\begin{point}{\revref}
     we can estimate amount of noise well
\end{point}

\reply{
    Added explanatory note aobut how to estimate noise.
}

\begin{point}{}
    The text at 468ff relating to the assumption of constant pop size and migration
    rates does a good job of explaining these consequences.  Perhaps all
    assumptions could be listed in a single place and their biological
    consequences explained.
\end{point}

\reply{
    The job of enumerating \emph{all} assumptions involved in modeling a real population
    with a system of discrete demes is daunting.
    We've put in more discussion of these pitfalls (as described above),
    but a comprehensive review is beyond the scope of this paper.
}

@@

\begin{point}{\revref}
    How exactly is the pairwise genetic distance calculated?  What does ``density
    of sites'' mean?  Is it just the fraction of sites at which two genomes
    differ?  If so it is just pairwise theta as used by Wakeley and others.  
    Eq 1:  what is $R_{ij}$ and how is $q$ calculated?
\end{point}

\reply{
    That's correct; we've reworded this to be more clear.
}

\begin{point}{}
    Eq \eqref{eq:commute_approx}: what is $R_{ij}$ and how is $q$ calculated?
\end{point}

\reply{
    We've added a clarifying sentence. \revref
}

@@

\begin{point}{}
    $G$, the movement rate, is clearly a key parameter but is not clearly defined (or
    if it was I missed it).  What are the units of G?  Geographic distance per
    unit time?  What are the time units?  Years?  Generations?  If the latter,
    then does G represent parent-offspring dispersal distance as defined by
    Wright?  Are generations assumed to be discrete?  If not, can a migrant
    leave offspring at various different geographic points?
\end{point}

\reply{
    A precise definition (in words and equations) is given where it is first introduced, at \revref.
    We've rewritten it to be hopefully more clear.
}

@@

\begin{point}{}
    As described, the model deals with groups of individuals (populations) that
    sometimes exchange individuals.  But much of landscape genetics/genomics
    deals with continuously-distributed species that don’t have recognizable
    populations.  Are the results also applicable to scenarios like this?  Why
    or why not?
\end{point}

\reply{
    This is an important point that applies not just to our method
    but to most methods of population genetics inference,
    and we view this disconnect as a major unsolved problem in the field.
    Our simulations of continuous space are explicitly designed
    to look at this problem.
    We've added substantially more discussion of this point
    in the interpretation of poplar results \llname{ll:continuous}.
}

\begin{point}{}
    The authors identify two sources of noise: Sampling noise (arising from
    sampling of individuals) and Process noise (stochasticity of population
    history).  Using SLiM, which I gather controls for random variation in the
    realized pedigree and hence eliminates (or only reduces?) Process noise,
    the authors obtain estimates for the relative magnitude of the two sources
    of noise.  This is useful but seems incomplete for two reasons:
    \begin{enumerate}
        \item       A third source of noise is sampling a finite number of
            genes or SNPs.  To some extent one can trade off sampling
            individuals vs loci, but there are limits to how effective that is.
        \item       The relative magnitudes of the three sources of error are
            not constant and can vary dramatically depending on parameter
            values.  Process error is inversely related to $N_e$ but otherwise
            fixed in magnitude; it does not change when sampling errors change.
            This means that Process error becomes relatively more important
            when there is a lot of replication in sampling.  In particular,
            genetic sampling error gets very small as the number of SNPs
            becomes arbitrarily large, in which case it is dwarfed by Process
            error (e.g., see Figure 6 in our 2009 Mol Ecol paper cited by the
            authors).
    \end{enumerate}
\end{point}

\reply{
    We certainly had omitted that source of noise,
    but have added it to the list now \revref.
    We agree it can be substantial, but aren't focusing on it here,
    as the contribution of sampling noise is much less appreciated,
    particularly in the context of continuous space (see our additional note on this).
    We're not sure what you mean about SLiM reducing process noise:
    on the contrary, since we use SLiM to simulate in continuous space,
    we expect process noise to be larger than in a randomly mating population.
}

@@

\begin{point}{}
    Also in relation to genomics-scale datasets, two general potential
    advantages can be identified:
    \begin{enumerate}
        \item       Arbitrarily large numbers of loci can increase precision
            and statistical power for existing analyses;
        \item       Ability to access information related to genome structure
            and function can provide novel insights and make possible
            qualitatively new types of analyses
    \end{enumerate}
    Lots of people have genomics-scale datasets these days so some discussion of
    this issue seems important.
\end{point}

\reply{
    Good point, but to keep the paper relatively focused we're not going into this in detail
    (but perhaps the pointer at {\revref} suffices?).
}

\begin{point}{}
    ``Can our method replace resistance-based methods? Perhaps, but the substantial
    uncertainty we saw on graphs with only tens of nodes is indicative of a
    larger problem we face for realistic models. We have seen that
    discretization of space results in substantial modeling error, in part
    because of randomness of geographic sampling and unmodeled process noise
    can lead to overfitting. Partitioning space into a finer grid should help
    with these problems, but tends to make the inference problem itself more
    ill-conditioned: with more connections, changing the value of one
    connection affects coalescence time less, and so inferences about that
    value must necessarily be less certain. This tradeoff implies some degree
    of unavoidable uncertainty.  Reproducible, reliable inference will likely
    require development of new inference methods that explicitly model
    continuous geography.''  This seems like a fairly crucial paragraph.  But
    what exactly is the take-home message for biologists interested in
    practical application of the methodology?
\end{point}

\reply{
    Good question!
    As written, the take-home was ``wait until we make better methods'',
    but we've added a better one. \revref
}

\begin{point}{}
    Coalescent rates are critical to model performance, which can be poor if the
    rate is too low.  It is obvious that the coalescent rate is inversely
    related to $N_e$, but in practical terms, what does it mean when the
    coalescent rate is lowered from 10 to 0.1?
\end{point}

\reply{
    We've added more context to this section. \revref
    % Assuming that the movement parameters are of order one, 
    % a coalescence rate of 10 means that 
    % lineages would be more likely to coalesce the first time they are in the same location
    % and a coalescence rate of 0.1 means that lineages would likely meet many times
    % before finally coalescing. 
    % For example, if the gene flow rates out of a deme for both lineages sum to 8
    % (rate of 1 for each direction for each lineage), 
    % the chance of the lineages coalescing before one of them leaves 
    % would be 10/18 (~56\%) with a coalescence rate of 10 
    % but only 0.1/8.1 (~1.2\%) with a coalescence rate of 0.1.
}

\begin{point}{\revref}
    ``we varied the standard deviation of the noise from 1/1000th to 1/50th of
    the mean value of $C$ for that grid.''  This sounds like a CV of 0.001 to
    0.02 for noise, both of which would be exceedingly small.  If that is not
    what is meant, please clarify.
\end{point}

\reply{
    That is what we mean, and although small, we don't think that's unreasonable for genome-scale data
    under a discrete population model (but see our later discussion of process noise).
    We realize that many people are working with much smaller number of (ascertained) loci,
    but there's enough bigger issues here related to continuous geography
    that we've not explored in that direction.
}

@@

\begin{point}{\revref}
    ``if fecundity is variable'' -- on what scale?  Among populations?  Among
    individuals?  Among ages within the same individual? 
\end{point}

\reply{
    We mean among individuals (and have added this).
}

\begin{point}{\revref}
    ``It is not required in this approach to have samples from every spatial
    location \ldots''  Required for what?  Are they saying that unsampled but real
    (ghost) populations have no effect on precision or bias?
\end{point}

\reply{
    We're simply saying that it is not a requirement to \emph{run} the method
    to have samples in each location --
    it is easy, in principle, to include ``ghost'' populations.
    Of course, unsampled populations will affect accuracy of the result,
    and we've added a note to clarify. \revref
    % In practice, the more locations with samples you have, the better the result will be.
    % We would not typically recommend doing inference that includes unsampled demes
    % if it could be avoided.
    % However, in the theoretical case of a population fitting the model perfectly
    % and knowing the other genetic distance values perfectly, 
    % it is possible to solve exactly for paramters relating to the unsampled populations
    % as long as there are as many unique equations as unknowns.
    % In a real situation, if data is very good for the sampled locations, 
    % inference should be reasonable for the unsampled ones.
}

@@

\begin{point}{}
    Many of the figures are too small to read easily without a magnifying glass.
    Figures 3-5 are large enough but they include miniscule fonts that make it
    impossible to tell exactly what is plotted.  Green and black do not provide
    enough contrast.
\end{point}

\reply{
    We sympathize, and have tried to improve this.
    However, some of this we think is unavoidable --
    the x-axis labels of figures like Figure~\ref{fig:5x3b_post_coalvcom}
    cannot be made larger without overlapping.
    Another approach would be to simply plot true versus observed values,
    which is equivalent to the current plot without the x-axis labels.
    We prefer to leave these labels, even if these are only really legible
    in the digital version (after zooming in).
}

@@

\begin{point}{\revref}
    another limitation of direct observations is that they only provide
    information about movement of individuals but not gene flow, which requires
    successful breeding
\end{point}

\reply{
    Good point, but we'd like to keep this bit more concise.
}

@@

\reply{
    Clarified. \revref
}

\begin{point}{\revref}
    need a semicolon, not a comma, after ``two genomes''
\end{point}

\reply{
    We reworded this.
}

\begin{point}{\revref}
    why quotes for ``from the model''?
\end{point}

\reply{
    Removed. \revref
}

### 2019-04-18 isolation_by_coalescence.tex
more responses

@@

The idea of ``resistance distance'' is an important tool in the landscape genomics toolbox.
Introduced by \citet{mcrae2006isolation},
it makes use of a mathematical equivalence between random walks and electrical resistance
\citep{nashwilliams1959random}:
the expected time for a continuous-time Markov chain that starts at node $x$
to first hit node $y$ and then return to $x$ (the ``commute time'')
is equal to the effective resistance between $x$ and $y$ in an electrical network
whose conductances are given by the movement rates of the Markov chain \citep{levin2008markov,doyle2006random}.
This measure therefore averages over all possible paths through the network between the two.
For this reason, 
if assigning local conductances based on the values of some landscape variable
results in a significant positive correlation between 
genetic distances between samples from different parts of the landscape
and the effective resistances between those locations,
then this is taken as good evidence that the landscape variable is a good indicator of
where gene flow occurs \citep{mcrae2007circuit,cushman2006complex}.
For instance,
if the conductance across a grid cell in a discretization of the landscape
is higher in flatter areas,
and this produces effective resistance values 
that are significantly correlated with genetic distances,
then one might conclude that the species in question tends to disperse more readily through flatter areas,
perhaps along river valleys.
More recent methods \citep{petkova2016visualizing,hanks2013circuit} seek to build \emph{de novo} 
a map of conductance values that produces resistances 
most strongly correlated with observed genetic distance,
and then interpret regions of low conductance in the resulting map as barriers to gene flow.

@@

In this paper, we contrast 
coalescence time and resistance distance,
and develop a method to infer movement rates on a discrete landscape
using genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry. \revpoint{3}{19}
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information; \revpoint{3}{20}
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem, \revpoint{3}{21}
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

The \emph{commute time} of a Markov chain between locations $i$ and $j$
is the expected time for a particle that moves according to the rules of the chain, 
started at $i$, to first reach $j$ and then return.
We will denote this by $R_{ij}$, to emphasize that it is also the effective resistance
between $i$ and $j$.
The \emph{coalescence time} is the expected time for two particles started at $i$ and $j$
to \emph{coalesce},
which happens at a given rate whenever they are in the same location.
We will denote this by $C_{ij}$.
Why should these be related to genetic distance? 
First, we explain this connection,
and then discuss their computation,
and finally develop a method that infers movement rates of the Markov chain
(which are reverse-time migration probabilities)
from observations of coalescence time. \revpoint{1}{9}

@@

There are a number of ways to calculate measures of dissimilarity using genetic data.
In this paper we use mean pairwise divergence, \revpoint{3}{6}
which for two genomes is calculated as the proportion of sites in the genome
at which the two genomes differ. \revpoint{3}{22}
It is more common in the literature to use $F_{ST}/(1-F_{ST})$, \revpoint{1}{7}
but the two are highly correlated for good theoretical reasons 
\citep{slatkin1991inbreeding,rousset1997genetic},
especially for SNP data.
(However, note that $F_{ST}$ depends on the sampling scheme, while divergence does not.) \revpoint{1}{7}
The two genomes differ at a site only if there has been a mutation in some ancestor on the lineages leading
from the two genomes back to their most recent common ancestor at that site.
Therefore,
under an infinite-sites model with average mutation rate $\mu$ per site and per meiosis,
pairwise divergence divided by $\mu$
is an unbiased estimate of the mean time to most recent common ancestor,
averaged across the genome \citep{hudson2007variance,ralph2015empirical}.
It is natural to model this process by following the two lineages back through time
(and across space)
until they find their most recent common ancestor,
i.e., until they coalesce.
Suppose we model the geographic distribution of the species
as a collection of randomly mating populations
that exchange occasional migrants,
thus discretizing geographic space.
If the genome is selectively neutral,
then the motion of these lineages back through time
forms a Markov process, the \emph{structured coalescent} \citep{wakeley2005coalescent}.
In this framework, each lineage performs a random walk across the populations,
with movement probabilities depending on the flux of individuals between populations.
This provides the link between genetic distance and random walks.

For data, we are given genotypes of individuals and their geographic coordinates;
then, we divide these individuals into groups,
and compute the genetic distance between group $i$ and group $j$
as the mean divergence between a pair of genomes, chosen one from each group.
We denote the resulting genetic distance $D_{ij}$;
since divergence is symmetric, $D_{ij} = D_{ji}$.
We compute $D_{ii}$ by sampling without replacement from group $i$,
so that generally $D_{ii} > 0$: the diagonal elements measure local diversity.
As discussed below, we will find commute times so that the mean commute time between
$i$ and $j$ matches $D_{ij}$ -- but the commute time of a location to itself is zero,
so we must account for local diversity.
We do this in the same way as \citet{petkova2016visualizing},
by introducing the \emph{local diversity} parameters $q_i$,
and using the model
\begin{align} \label{eq:commute_approx}
	D_{ij} \approx \comdist_{ij} = R_{ij}/4 + q_{i}/2 + q_{j}/2 .
\end{align}
The factor of four is necessary to make commute time work on the same time scale as coalescence time:
the commute time covers the path from $i$ to $j$ twice
(once in both directions),
while the coalescence time covers the distance only once,
and in half the time, because two particles are moving.
% Furthermore, coalescence time includes within-location diversity
% because particles do not coalesce instantly --
% therefore, commute time is most analogous to the first meeting time.
For a conceptual picture of the differences between the coalescence time and commute time,
see Figure \ref{fig:concept_coalcom}.

@@

Now, we explain how to compute the relevant quantities --
mean commute and coalescence times -- from a Markov chain model of lineage movement.
The Markov chain is defined by its \emph{generator matrix}, denoted $G$,
for which $G_{xy}$ gives the jump rate from $x$ to $y$.
This means that if the Markov chain is at location $X_t$ at time $t$,
then for $x \neq y$, the probability the chain jumps from $x$ to $y$ in the next unit of time
is proportional to $G_{xy}$,
i.e., $\PP{X_{t+\epsilon} = y \given X_t = y} = \epsilon G_{xy} + O(\epsilon^2)$,
and $G_{xx} = - \sum_{y \neq x} G_{xy}$.
We will need to find the \emph{hitting times} of the chain --
i.e., for each pair of states $x$ and $y$, 
the mean time until the chain first hits $y$ after being started at $x$.
We denote this quantity $H_{xy}$.
Throughout, we assume that all hitting times are finite,
which implies the chain is connected and irreducible.
By conditioning on $X_\epsilon$,
one can use the Markov property to show that
\begin{align} \label{eqn:hitting_sum}
    \sum_y G_{xy} H_{yz} = -1 \qquad \text{for} \; x \neq z,
\end{align}
i.e., if $G_{-z}$ is the matrix with the $z^\text{th}$ row and column removed,
and $\bone$ is the vector composed of all ones,
then the $z^\text{th}$ column of $H$, except for $H_{zz}$,
can be computed as $- (G_{-z})^{-1} \bone$.
This, along with $H_{zz} = 0$, allows computation of $H$
from the movement rates $G$ \citep{kemeny1983finite}.

>>>>

\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}% 
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

@@

The idea of ``resistance distance'' is an important tool in the landscape genomics toolbox.
Introduced by \citet{mcrae2006isolation},
it makes use of a mathematical equivalence between random walks and electrical resistance
\citep{nashwilliams1959random}:
the expected time for a continuous-time Markov chain that starts at node $x$
to first hit node $y$ and then return to $x$ (the ``commute time'')
is equal to the effective resistance between $x$ and $y$ in an electrical network
whose conductances are given by the movement rates of the Markov chain \citep{levin2008markov,doyle2006random}.
This measure therefore averages over all possible paths through the network between the two.
Suppose we
assign local conductances based on the values of some landscape variable, 
and compute the correlation of
genetic distances between samples from different parts of the landscape
with the effective resistances between the samples' locations.
\revpoint{4}{2}
If this results in a significant positive correlation,
then this is taken as good evidence that the landscape variable is a good indicator of
where gene flow occurs \citep{mcrae2007circuit,cushman2006complex}.
For instance,
if the conductance across a grid cell in a discretization of the landscape
is higher in flatter areas,
and this produces effective resistance values 
that are significantly correlated with genetic distances,
then one might conclude that the species in question tends to disperse more readily through flatter areas,
perhaps along river valleys.
More recent methods \citep{petkova2016visualizing,hanks2013circuit} seek to build \emph{de novo} 
a map of conductance values that produces resistances 
most strongly correlated with observed genetic distance,
and then interpret regions of low conductance in the resulting map as barriers to gene flow.

@@

In this paper, we contrast 
coalescence time and resistance distance,
and develop a method to infer movement rates on a discrete landscape
using genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may infer movement rates uncorrelated with the truth \revpoint{3}{19}
if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry. 
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep[``Estimating Effective Migration Surfaces'',][]{petkova2016visualizing} infer conductances without such prior information; \revpoint{3}{20}
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this coarse resolution reflects a more general problem, \revpoint{3}{21}
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

The \emph{commute time} of a Markov chain between locations $i$ and $j$
is the expected time for a particle that moves according to the rules of the chain, 
started at $i$, to first reach $j$ and then return.
We will denote this by $R_{ij}$, to emphasize that it is also the effective resistance
between $i$ and $j$.
The \emph{coalescence time} is the expected time for two particles started at $i$ and $j$
to \emph{coalesce},
which happens at a given rate whenever they are in the same location.
We will denote this by $C_{ij}$.
Why should these be related to genetic distance? 
First, we explain this connection,
and then discuss their computation,
and finally develop a method that infers movement rates of the Markov chain
(which are reverse-time migration probabilities)
from observations of coalescence time, \revpoint{1}{9}
and apply this to a real dataset.
Although we have tried to highlight the intuitive aspects of this initial technical discussion,
readers who are interested in practical differences
more than the mathematical arguments should be able to skip to ``Test cases and simulations''
without loss of continuity. \llabel{ll:flagging}

@@

There are a number of ways to calculate measures of dissimilarity using genetic data.
In this paper we use mean pairwise divergence, \revpoint{3}{6}
which for two genomes is calculated as the proportion of sites in the genome
at which the two genomes differ. \revpoint{3}{22}
It is more common in the literature to use $F_{ST}/(1-F_{ST})$, \revpoint{1}{7}
but the two are highly correlated for good theoretical reasons 
\citep{slatkin1991inbreeding,rousset1997genetic},
especially for SNP data.
(However, note that $F_{ST}$ depends on the sampling scheme, while divergence does not.) \revpoint{1}{7}
The two genomes differ at a site only if there has been a mutation in some ancestor on the lineages leading
from the two genomes back to their most recent common ancestor at that site.
Therefore,
under an infinite-sites model with average mutation rate $\mu$ per site and per meiosis,
pairwise divergence divided by $\mu$
is an unbiased estimate of the mean time to most recent common ancestor,
averaged across the genome \citep{hudson2007variance,ralph2015empirical}.
It is natural to model this process by following the two lineages back through time
(and across space)
until they find their most recent common ancestor,
i.e., until they coalesce.
Suppose we model the geographic distribution of the species
as a collection of randomly mating populations
that exchange occasional migrants,
thus discretizing geographic space.
It is common to assume that
the motion of these lineages back through time
forms a Markov process, the \emph{structured coalescent} \citep{wakeley2005coalescent}.
(This assumption requires the effects of natural selection on each segment of the genome 
to be generally weak.) \revpoint{3}{1}
In this framework, each lineage performs a random walk across the populations,
with movement probabilities depending on the flux of individuals between populations.
This provides the link between genetic distance and random walks.

For data, we are given genotypes of individuals and their geographic coordinates;
then, we divide these individuals into groups,
and compute the genetic distance between group $i$ and group $j$
as the mean divergence between a pair of genomes, chosen one from each group.
We denote the resulting genetic distance $D_{ij}$;
since divergence is symmetric, $D_{ij} = D_{ji}$.
We compute $D_{ii}$ by sampling without replacement from group $i$,
so that generally $D_{ii} > 0$: the diagonal elements measure local diversity.
As discussed below, we will find commute times so that the mean commute time between
$i$ and $j$ matches $D_{ij}$ -- but the commute time of a location to itself is zero,
so we must account for local diversity.
We do this in the same way as \citet{petkova2016visualizing},
by introducing the \emph{local diversity} parameters $q_i$,
and using the model
\begin{align} \label{eq:commute_approx}
	D_{ij} \approx \comdist_{ij} = R_{ij}/4 + q_{i}/2 + q_{j}/2 .
\end{align}
Concretely, this means finding 
movement rates of a Markov chain whose matrix of commute times ($R$)
yields a good approximation to the matrix of genetic distances,
after adding local diversity parameters ($q$, which are also free parameters). \revpoint{3}{7}
The factor of four is necessary to make commute time work on the same time scale as coalescence time:
the commute time covers the path from $i$ to $j$ twice
(once in both directions),
while the coalescence time covers the distance only once,
and in half the time, because two particles are moving.
% Furthermore, coalescence time includes within-location diversity
% because particles do not coalesce instantly --
% therefore, commute time is most analogous to the first meeting time.
For a conceptual picture of the differences between the coalescence time and commute time,
see Figure \ref{fig:concept_coalcom}.

A simple situation when the two clearly differ \llabel{ll:threepop}
is in the presence of biased gene flow,
as for instance between a source and a sink population:
coalescence only requires that lineages move in \emph{one} direction
between populations,
while a commute requires movement in both.

@@

Now, we explain how to compute the relevant quantities --
mean commute and coalescence times -- from a Markov chain model of lineage movement.
The Markov chain is defined by its \emph{generator matrix}, denoted $G$,
for which $G_{xy}$ gives the ``jump rate'' from $x$ to $y$.
The jump rate $G_{xy}$ is \revpoint{3}{8}
the probability the chain jumps to position $y$ when at position $x$ per unit of time.
(More precisely, if the Markov chain is at location $X_t$ at time $t$, 
then for $x \neq y$, $\PP{X_{t+\epsilon} = y \given X_t = x} = \epsilon G_{xy} + O(\epsilon^2)$.)
For mathematical convenience, the diagonal of this matrix is completed so that rows sum to zero:
$G_{xx} = - \sum_{y \neq x} G_{xy}$.
We will need to find the \emph{hitting times} of the chain --
i.e., for each pair of states $x$ and $y$, 
the mean time until the chain first hits $y$ after being started at $x$.
We denote this quantity $H_{xy}$.
Throughout, we assume that all hitting times are finite,
which implies the chain is connected and irreducible.
By conditioning on $X_\epsilon$,
one can use the Markov property to show that
\begin{align} \label{eqn:hitting_sum}
    \sum_y G_{xy} H_{yz} = -1 \qquad \text{for} \; x \neq z,
\end{align}
i.e., if $G_{-z}$ is the matrix with the $z^\text{th}$ row and column removed,
and $\bone$ is the vector composed of all ones,
then the $z^\text{th}$ column of $H$, except for $H_{zz}$,
can be computed as $- (G_{-z})^{-1} \bone$.
This, along with $H_{zz} = 0$, allows computation of $H$
from the movement rates $G$ \citep{kemeny1983finite}.
